---
atitle: CourseProject1_ReproductibleAnalysis
author: "Mihaela"
date: "4/21/2021"
output:
  md_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Loading Required packages

```{r}
library(dplyr)
library(ggplot2)
```

## Downloading and Processing the data
```{r}
url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
dest <- "C:/Users/Mihaela/AppData/Roaming/SPB_16.6/R-test/datasciencecoursera/Explanatory_Data_Analysis/activity.zip"
download.file(url,dest)
data <- unz(dest,filename = "activity.csv")
```

Loading the csv file 
Once we load the file, we will compose two data frames - one with NA (df.na) and one without NA(df)

```{r}
df <- read.table(data,
                 header = T,
                 sep = ",",
                 na.strings = "NA")
df.na <- df
df <- na.omit(df)
```

Using dplyr to sum the steps for each day and convert date to a date format
```{r}
activity <- df %>% group_by(date) %>% summarise(steps = sum(steps)) %>% 
  mutate(date = as.Date(date, format = "%Y-%m-%d"))
```

## Histogram of the total number of steps taken each day and summary of the data
```{r}
hist(activity$steps, main = "total number of steps taken each day", xlab = "steps")
summary(activity)
```
## What is the average daily activity pattern?
```{r}
activity_interval <- aggregate(steps ~ interval, df, mean)

plot(activity_interval$interval,activity_interval$steps, type = "l",
     main = "average daily activity pattern", xlab = "interval", ylab = "steps",
     col = "blue")
```
Identifying the interval with maximum steps
```{r}
activity_interval[which.max(activity_interval$steps),]
```
## Imputing missing values. 
I have used dplyr to make new column with the data of the total steps each day. And then replaced the NA with the average steps. for a day.
```{r}
sum(is.na(df.na))
df_imputed <-df.na %>% 
             mutate(date = as.Date(date, format = "%Y-%m-%d")) %>%
             group_by(date) %>%
             mutate(sum_steps = sum(steps))

df_imputed$sum_steps[which(is.na(df_imputed$sum_steps))] = mean(df_imputed$sum_steps,na.rm = T)
```  
Making a new histogram with the imputed data frame
```{r}
hist(df_imputed$sum_steps, main = "Total Number of Steps Taken Each Day", xlab = "Steps", breaks = 30)
```

## Are there differences in activity patterns between weekdays and weekends?
Preparing the data

The first thing we want to do is use the weekend function to transform the dates to the names of each day.
Then we will replace each day with the corresponding label weekday or weekend.
```{r}
df_imputed$days <- weekdays(df_imputed$date)
df_imputed$days <- ifelse(df_imputed$days=="Saturday" | df_imputed$days=="Sunday",
                          "weekend","weekday")
```
Because in the last example we imputed the data with the average of the day. Now we have to impute using the average of each interval.
```{r}
df_imputed_intervals <-aggregate(steps~interval+days,data=df_imputed,FUN=mean,na.action=na.omit)
```
Plotting the result
```{r}
g <- ggplot(df_imputed_intervals, aes(interval,steps))
g + facet_wrap(~days) + geom_line() 
```
